---
title: "TEOS Hydrostation S Data Lab"
author: "Raymond U. Infante"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
Link for information on BATS Hydrostation S
https://bats.bios.edu/about/



## Requiered Libraries
```{r, message=FALSE}
library(tidyverse)
#Install (seacarb)
#Gibbs Seawater
library(gsw)
```


## Import the data
```{r message=FALSE, warning=FALSE}
hydrostation_bottle <- read_delim("hydrostation_bottle.txt", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 31)



hydrostation_bottle_names <- read_csv("hydrostation_bottle.txt", 
    skip = 30)

#Fix names for the dataset
colnames(hydrostation_bottle) = colnames(hydrostation_bottle_names)

view(hydrostation_bottle)

```
## Variables Names and Units
- yyyymmdd = Year Month Day   
- decy   = Decimal Year     
- time   = Time (hhmm)      
- latN   = Latitude (Deg N) 
- lonW   = Longitude (Deg W)
- Depth  = Depth (m)                  
- Temp   = Temperature ITS-90 (C) 
- Pres   = CTD Pressure (dbar)   
- CTD_S  = CTD Salinity (PSS-78)      
- Sal1   = Salinity-1 (PSS-78)        
- Sig-th = Sigma-Theta (kg/m^3) - density of seawater at a given temperature (Density - 1000)   
- O2(1)  = Oxygen-1 (umol/kg)          
- OxFixT = Oxygen Fix Temp (C)        
- Anom1  = Oxy Anomaly-1 (umol/kg)    

## Quality flags
*  -999  = No data
*  0 = Less than detection limit

```{r}
hydrostation_bottle %>% 
  filter(`Sig-th` != -999 & Depth < 20) %>% 
  ggplot()+
  geom_point(aes(x = decy, y = `Sig-th`))+
  geom_line(aes(x =decy, y = `Sig-th` ))
#Clear Seasonal signal for sigma-theta
hydrostation_bottle %>% 
  filter(`Sig-th` != -999 & Depth < 20) %>% 
  ggplot()+
  geom_point(aes(x =Temp, y = `Sig-th` ))

#Density and Temperature are strongly correlated, but there appears to be two outliners that we will likely need to address at some point

#We only have density data from 1988-present, but temperature and salinity data from 1950s-present, this mean i can calculate seawater density from 1950s- present, by using TEOS-10. 
```

## TEOS-10 Toolbox in package seacarb

```{r, include=FALSE, echo=FALSE}
#?gsw
#?gsw_sigma0 # It said we need absolute salinity and conservative temperature (SA (g/kg),CT (°C))
#?gsw_SA_from_SP #Absolute Salinity from Practical Salinity
#Practical Salinity*
#Sea pressure (dbar)
#Longitude
#Latitude

hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = Pres ))
#Not for all the time series. Missing from 1980 

hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = Depth ))
#We have Depth for the entire time series

#?gsw_p_from_z()

hydrostation_bottle =  hydrostation_bottle %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN))

hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = Pres, y = Pres_gsw))
# We see strong 1:1 agreement between measured pressure and calculated pressure

#Checking lat, lon, and salinity data
hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = CTD_S))

hydrostation_bottle =  hydrostation_bottle %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN)) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(Sal1, Pres_gsw,360-lonW, latN))

#Checking the Absolute salinity
hydrostation_bottle %>% 
  ggplot()+
  geom_point(aes(x = decy, y = s_abs_gsw))
#Plotting the ratio for Absoute Salinity (created) vs Sal1 (Given in the dataset)
hydrostation_bottle %>% 
  filter(Sal1!= -999) %>% 
  ggplot()+
  geom_point(aes(x = Sal1, y = s_abs_gsw))

#Temperature to Conservative Temperature
#?gsw_CT_from_t


HydroS =  hydrostation_bottle %>% 
  filter(Sal1!= -999) %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN)) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(Sal1, Pres_gsw,360-lonW, latN)) %>% 
  mutate(T_Cons_gsw = gsw_CT_from_t(s_abs_gsw,Temp,Pres_gsw))

HydroS %>% 
  filter(Temp != -999) %>% 
  ggplot()+
  geom_point(aes(x = Temp, y = T_Cons_gsw))




HydroS =  hydrostation_bottle %>% 
  filter(Sal1 != -999 & `Sig-th` != -999) %>% 
  mutate(Pres_gsw = gsw_p_from_z(Depth*-1, latN)) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(Sal1, Pres_gsw,360-lonW, latN)) %>% 
  mutate(T_Cons_gsw = gsw_CT_from_t(s_abs_gsw,Temp,Pres_gsw)) %>% 
  mutate(sig_theta_gsw = gsw_sigma0(s_abs_gsw,T_Cons_gsw))

HydroS %>% 
  filter(`Sig-th` != -999) %>% 
  ggplot()+
  geom_point(aes(x = `Sig-th`, y = sig_theta_gsw))

HydroS %>% 
  filter(sig_theta_gsw < 0) %>% 
  view()

HydroS_Corrected_a = HydroS %>% 
  filter(sig_theta_gsw<0) %>% 
  mutate(s_abs_gsw = gsw_SA_from_SP(CTD_S, Pres_gsw,360-lonW, latN)) %>% 
  mutate(T_Cons_gsw = gsw_CT_from_t(s_abs_gsw,Temp,Pres_gsw)) %>% 
  mutate(sig_theta_gsw = gsw_sigma0(s_abs_gsw,T_Cons_gsw)) 
  

HydroS_Corrected_b = HydroS %>% 
  filter(sig_theta_gsw>0) 

#The corrected one. And last one
HydroS_corrected = rbind(HydroS_Corrected_a,HydroS_Corrected_b) 



                                           
```

For Homework: 

try to replace the broken line of data, and make in a way the code use in 161-168 in only a pipeline. 
replace a line in the data using the tidyverse.





```{r}
view(HydroS)

HydroS %>% 
  mutate(test = if_else(sig_theta_gsw<0, mutate(s_abs_gsw = gsw_SA_from_SP(CTD_S, Pres_gsw,360-lonW, latN)) %>% 
  mutate(T_Cons_gsw = gsw_CT_from_t(s_abs_gsw,Temp,Pres_gsw)) %>% 
  mutate(sig_theta_gsw = gsw_sigma0(s_abs_gsw,T_Cons_gsw)) , "Hi"))


```

