---
title: "TEOS Hydrostation S Data Lab"
author: "Raymond U. Infante"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## Requiered Libraries
```{r, message=FALSE}
library(tidyverse)
#Install (seacarb)
library(gsw)
```


## Import the data
```{r message=FALSE, warning=FALSE}
hydrostation_bottle <- read_delim("hydrostation_bottle.txt", 
    delim = "\t", escape_double = FALSE, 
    col_names = FALSE, trim_ws = TRUE, skip = 31)



hydrostation_bottle_names <- read_csv("hydrostation_bottle.txt", 
    skip = 30)

#Create names for the dataset
colnames(hydrostation_bottle) = colnames(hydrostation_bottle_names)

```
## Variables Names and Units
- yyyymmdd = Year Month Day   
- decy   = Decimal Year     
- time   = Time (hhmm)      
- latN   = Latitude (Deg N) 
- lonW   = Longitude (Deg W)
- Depth  = Depth (m)                  
- Temp   = Temperature ITS-90 (C) 
- Pres   = CTD Pressure (dbar)   
- CTD_S  = CTD Salinity (PSS-78)      
- Sal1   = Salinity-1 (PSS-78)        
- Sig-th = Sigma-Theta (kg/m^3) - density of seawater at a given temperature (Density - 1000)   
- O2(1)  = Oxygen-1 (umol/kg)          
- OxFixT = Oxygen Fix Temp (C)        
- Anom1  = Oxy Anomaly-1 (umol/kg)    

## Quality flags
*  -999  = No data
*  0 = Less than detection limit

```{r}
hydrostation_bottle %>% 
  filter(`Sig-th` != -999 & Depth < 20) %>% 
  ggplot()+
  geom_point(aes(x = decy, y = `Sig-th`))+
  geom_line(aes(x =decy, y = `Sig-th` ))
#Clear Seasonal signal for sigma-theta
hydrostation_bottle %>% 
  filter(`Sig-th` != -999 & Depth < 20) %>% 
  ggplot()+
  geom_point(aes(x =Temp, y = `Sig-th` ))

#Density and Temperature are strongly correlated, but there appears to be two outliners that we will likely need to address at some point

#We only have density data from 1988-present, but temperature and salinity data from 1950s-present, this mean i can calculate seawater density from 1950s- present, by using TEOS-10. 
```


